---
title: "So Many Ads, So Many Themes"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(highcharter)
library(gapminder)
library(htmlwidgets)
library(kableExtra)
library(fmsb)
library(plotly)
```

```{r upload files, echo=FALSE, message=FALSE}
data_ad_time <- read_csv("~/STA 518/Final Project/Portfolio/Final Proposal/Final Project/Ad  Time _ Category - Sheet1.csv")
data <- read_csv("~/STA 518/Final Project/Portfolio/Final Proposal/Final Project/combined  - Sheet1.csv")
data_ad_cost <- read_csv("~/STA 518/Final Project/Portfolio/Final Proposal/Final Project/Cost per Super bowl ad 2000-2020 - Sheet1 (1).csv")
```
```{r joining two data sets, echo=FALSE, message=FALSE, results="hide"}
# merge data and data_ad_cost together 
total <- inner_join(data, data_ad_cost, by="year")
total
dim(total)
knitr::kable(total)
#chr to Numeric cost_per_minute and cost_per_thirty_seconds
total11<-
total%>% as.data.frame %>%
  mutate(cost_per_minute=str_remove_all(cost_per_minute,"[$,]"),
         cost_per_minute=as.numeric(cost_per_minute)) %>%
  mutate(cost_per_thirty_seconds=str_remove_all(cost_per_thirty_seconds,"[$,]"),
        cost_per_thirty_seconds=as.numeric(cost_per_thirty_seconds))%>%
  
  mutate(cost_per_second = cost_per_minute/60 )%>%
  
  mutate(cost_of_ad= cost_per_second*ad_time) %>%
  mutate(cost_of_ad_in_millions =cost_of_ad/1000000 )%>%
  mutate(cost_per_second_in_thousands =cost_per_second/100000)
  
total11
 # mutate(currency = scales::dollar(cost_per_second))
  
 
  
```

```{r exploratory bar charts, echo=FALSE, message=FALSE}
# count of brands use of gender
ggplot(data=total, aes(x=category)) +
  geom_bar(stat= "count", position="dodge",color= "black", aes(fill=celebrity))+ 
   ggtitle("Super Bowl Ad Categories and Celebrity Apperances From 2000-2020")
```
This graph shows celebrity appearances for each category from 2000-2020. As we can see the beer brands did not use very many celebrities to promote their drink. Although the sample size of ads was heavily favored in the beer category we can see that the other categories followed the same trend. Not as drastically, but it does look like celebrities were not getting their fame from super bowl ads from the year 2000-2020.


```{r porportions, echo=FALSE, message=FALSE}
 ggplot(data=total11, aes(x= year, fill=use_sex))+
   geom_bar(stat= "count", position="fill")+
   ggtitle("Porportion of SuperBowl Ads that use Sex Appeal from 2000 to 2020")
```

This graph depicts the use of 'Sex Appeal' in all brands studied in this sample. In the year 2000-2009 it looks like brands were using this to attract consumers as this number of ads was hovering around the 50% mark. From then on it has drastically decreased as brands have found other ways of promoting their products. It seems that sex appeal does not have much consumer pull as it once had in the early 2000's. 

```{r ad_themes, echo=FALSE, message=FALSE}
#creating frequency table 
 
total11 %>%
  group_by(ad_time) %>%
  summarize(Frequency=n()) %>%
  arrange(desc(Frequency)) %>%
  mutate(Proportion=Frequency/sum(Frequency)) %>%
  mutate(AdTime= ad_time)%>%
  select(AdTime, Frequency, Proportion)%>%
  print(n=50)
```
  

```{r percentage, echo=FALSE, message=FALSE}
 total12<- total11 %>%
  group_by(animals) %>%
  summarise(counts=n(),
            percentage= n()/nrow(total11))
 #pie chart
 ggplot(data=total12, aes(x="", y=percentage, fill=animals))+
  geom_col(color="blue")+
  coord_polar("y", start=0)+
  geom_text(aes(label= paste(round(percentage*100), "% ")),
            position = position_stack(vjust = 0.5))+
  theme(panel.background = element_blank(),
        axis.line= element_blank(),
         axis.text= element_blank(),
         axis.ticks= element_blank(),
        plot.title=element_text(hjust=0.5, size=18))+
  ggtitle("Pie Chat of Percentage of Ads that use Animals")+
   scale_fill_brewer("Use Animals")

```
For years 2000-2020 it seems that animal use is not as frequent as one may think. Although everyone loves the Budweiser horses and puppy, not many other brands are using animals to promote their product. 

```{r pie chartf, message=FALSE, echo=FALSE}
 #percentage
 total12<- total11 %>%
  group_by(funny) %>%
  summarise(counts=n(),
            percentage= n()/nrow(total11))
 #pie chart humor
 ggplot(data=total12, aes(x="", y=percentage, fill=funny))+
  geom_col(color="blue")+
  coord_polar("y", start=0)+
  geom_text(aes(label= paste0(round(percentage*100), "% ")),
            position = position_stack(vjust = 0.5))+
  theme(panel.background = element_blank(),
        axis.line= element_blank(),
         axis.text= element_blank(),
         axis.ticks= element_blank(),
        plot.title=element_text(hjust=0.5, size=18))+
  ggtitle("Pie Chat of Percentage of Ads that use humor")+
   scale_fill_brewer("Use Humor") 

```
From the year 2000-2020 it seems that super bowl ads have used humor in most of their ads. With almost 70% of the ads sampled, humor is included. From personal experience it does seem humor has found its way into a lot of ads, even outside of super bowl viewing. It will be interesting to see if this trend continues or if there will be another type of ad that takes over.

```{r pie chart, message=FALSE, echo=FALSE}

 #percentage
 total12<- total11 %>%
  group_by(danger) %>%
  summarise(counts=n(),
            percentage= n()/nrow(total11))
 #pie chart danger
 ggplot(data=total12, aes(x="", y=percentage, fill=danger))+
  geom_col(color="blue")+
  coord_polar("y", start=0)+
  geom_text(aes(label= paste0(round(percentage*100), "% ")),
            position = position_stack(vjust = 0.5))+
  theme(panel.background = element_blank(),
        axis.line= element_blank(),
         axis.text= element_blank(),
         axis.ticks= element_blank(),
        plot.title=element_text(hjust=0.5, size=18))+
  ggtitle("Pie Chat of Percentage of Ads that use Danger")+
   scale_fill_brewer("Use Danger") 

```
