---
title: "How Much has Ad Price Gone Up, You Ask?!"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(readr)
library(tidyverse)
library(highcharter)
library(gapminder)
library(htmlwidgets)
library(kableExtra)
library(fmsb)
library(plotly)
library(ggcorrplot)
```

```{r upload files, echo=FALSE, message=FALSE}
data_ad_time <- read_csv("~/STA 518/Final Project/Portfolio/Final Proposal/Final Project/Ad  Time _ Category - Sheet1.csv")
data <- read_csv("~/STA 518/Final Project/Portfolio/Final Proposal/Final Project/combined  - Sheet1 (3).csv")
data_ad_cost <- read_csv("~/STA 518/Final Project/Portfolio/Final Proposal/Final Project/Cost per Super bowl ad 2000-2020 - Sheet1 (1).csv")
```

```{r joining two data sets, echo=FALSE, message=FALSE, results="hide"}
# merge data and data_ad_cost together 
total <- inner_join(data, data_ad_cost, by="year")
total
dim(total)
knitr::kable(total)
#chr to Numeric cost_per_minute and cost_per_thirty_seconds
total11<-
total%>% as.data.frame %>%
  mutate(cost_per_minute=str_remove_all(cost_per_minute,"[$,]"),
         cost_per_minute=as.numeric(cost_per_minute)) %>%
  mutate(cost_per_thirty_seconds=str_remove_all(cost_per_thirty_seconds,"[$,]"),
        cost_per_thirty_seconds=as.numeric(cost_per_thirty_seconds))%>%
  
  mutate(cost_per_second = cost_per_minute/60 )%>%
  
  mutate(cost_of_ad= cost_per_second*ad_time) %>%
  mutate(cost_of_ad_in_millions =cost_of_ad/1000000 )%>%
  mutate(cost_per_second_in_thousands =cost_per_second/100000)
  
total11
 # mutate(currency = scales::dollar(cost_per_second))
  
 
  
```
```{r }
#total1<- bind_cols(data_ad, data_ad_time %>% select(-year, -brand))
```


```{r linegraph}
total3<- total11%>% group_by(year)%>%
summarise(avg_ad_cost= mean(cost_of_ad)) 

hc <- total3 %>% 
  hchart('line', hcaes(x = year, y =avg_ad_cost, group = year))%>%
  hc_title(text="Average Cost of Super Bowl Ads Over 20 Years")
  
  hc
```
From 2000 to 2020, the cost of Super Bowl ads has been steadily increasing starting at around 2.2 million in 2000 to 11.9 million in 2020.  The average price only slightly decreased from 2001-2002, 2003-2004, 2014-2015, and from 2017-2018. From 2000 to 2020, the average Super Bowl ad cost had increased by more than 500%.

```{r boxplot, reorder }
#reordering boxplots based on median cost_of_ad
ggplot(data=total11, aes(x= fct_reorder(category, cost_of_ad_in_millions), y=cost_of_ad_in_millions))+
geom_boxplot(fill="lightblue")+
ggtitle("Boxplots of Cost Per Ad by Category")+
xlab("Ad Category")+
ylab("Ad Cost in Millions")
```
From these boxplots, we can see that the NFL and Car ads look to have significantly larger median ad costs. The Finance, Beer, Snack, and Pop categories have more similar median ad costs.
 
```{r sunburst}
total3 <- data_to_hierarchical(data=total11, c(category, brand),cost_of_ad_in_millions)


hchart(total3, type = "sunburst")%>%
hc_title(text="Total Ad Cost by Category and Brand in Millions from 2000-2020")

hchart(total3, type = "treemap")
```
From the Sunburst plot and square plot, we can see the ad categories broken down into brands. When you hover the chart, it shows you how much money in millions that the brand/category has spent on Super Bowl ads in the past two decades. 

```{r bubble plot}
p<- ggplot(data=total11, aes(x=year, y=cost_of_ad_in_millions, fill= category))+
  geom_point(alpha=0.5, shape=21, size=3)+
  ylab("Cost of Ad in Millions")+
  xlab("Year")+
  ggtitle("Ad Cost by Year and Category")
  
 ggplotly(p, tooltip="text")


```
This bubble plot breaks down ad cost by category. We can see that the NFL ads cost significantly more in the early 2000’s, late 2010’s, and in 2020. This is because the NFL ads are longer than most of the other ads. Although, based on our research, the NFL likely doesn’t actually pay this much for their own ads. From this plot, we can also see that car ads are more expensive and became more frequent in the 2010’s. 

```{r coorelation, echo=FALSE, message=FALSE}
#Correlation Matrix
total3<- total11%>%
select(ad_time, cost_of_ad, year, cost_per_second)
p.mat<- cor_pmat(total3)
corr<- round(cor(total3), 1)
ggcorrplot(corr, hc.order =TRUE,
           type="lower", p.mat=p.mat,
           title= "Correlation of Super Bowl Ad Times, Cost, and Year",
           legend.title= "Pearson Correlation", lab=TRUE, lab_col="white", lab_size=2 , method="square",
           ggtheme= theme_grey, outline.color="black")

```
From this correlation matrix, we can see that cost_per_second and year have a perfect correlation. That’s because the price per second has increased from year to year. Additionally, we see that ad_time and cost_of_ad are very highly correlated since the longer the ad, the more expensive it is.